Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828

### 2017-02-14 inst/examples/20170201-manuscript.Rmd
updateing analyses

```{r 201606krakow-init2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Full combined catalogue (Fen + Kun) with marked duplicates
df.combined.preprocessed <- readRDS("df.combined.Rds")
# Data with duplicates removed
df.full <- df.combined.preprocessed %>% filter(!remove)
# Data with years limited as well
df0 <- df.full %>% filter(publication_year >= min.year & publication_year <= max.year)
# Calculate title length (nchar and word count)
df0$title_length_nchar <- nchar(df0$title)
df0$title_length_wordcount <- sapply(strsplit(df0$title, " "), length)
df0$title_length_wordcount[is.na(df0$title)] <- NA
```

@@

```{r places_per_decade, message=FALSE, warning=FALSE, fig.width=10, fig.height=4, echo=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw(20))
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>%
              group_by(catalog, publication_decade) %>%
	      summarise(n = length(unique(na.omit(publication_place))))
p <- ggplot(df, aes(x = publication_decade, y = n, shape = catalog, linetype = catalog)) + 
       geom_point(size = 4) + 
       geom_line(size = 1) + 
       ylab("Places (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen")) + 
       guides(shape = guide_legend(title = "Catalog", reverse = TRUE)) +
       guides(linetype = guide_legend(title = "Catalog", reverse = TRUE)) +       
       ggtitle("Unique publication places")
print(p)
```

@@

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=4, echo=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw(20))
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>% group_by(catalog, publication_decade) %>% tally()
p <- ggplot(df, aes(x = publication_decade, y = n, shape = catalog, linetype = catalog)) + 
       geom_point(size = 4) + 
       geom_line(size = 1) + 
       ylab("Documents (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen")) +
       guides(shape = guide_legend(title = "Catalog", reverse = TRUE)) +
       guides(linetype = guide_legend(title = "Catalog", reverse = TRUE)) +  
       ggtitle("Title count")
print(p)
```

@@

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=9, fig.height=5, fig.show="hold", out.width="400px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n())

  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 20)), "NA"))
  df2$highlight <- rep("Other", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "Octavo"
  df2$gatherings <- factor(df2$gatherings, levels = levels(df0$gatherings))

@@

```{r title_length_by_lang, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
dfo <- df0

# Analysis time window
dfo <- dfo %>% filter(publication_year >= 1640 & publication_year <= 1911)

top.language <- setdiff(names(top(dfo, "language", 11)), "NA")

df <- dfo %>% filter(language %in% top.language) %>%
              group_by(publication_decade, language) %>%
              summarize(mean_title_length = mean(title_length_wordcount),
	      		n = n())

p <- ggplot(df, aes(x = publication_decade, y = mean_title_length)) +
       geom_point(aes(size = n, shape = language), size = 5) +
       geom_line(aes(linetype = language), size = 1) +
       guides(
            shape = guide_legend(title = "Language"),
            linetype = guide_legend(title = "Language")
	    ) +
       ylab("Title length (word count)") +
       xlab("Publication decade") +
       ggtitle("Title length")

@@

```{r comparisons, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=18, fig.height=10}
pics <- list()
for (catalogue in na.omit(unique(df0$catalog))) {

@@

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=4}
library(magrittr)
library(ggplot2)
library(dplyr)
v <- seq(1500, 1800, 100)
theme_set(theme_bw(20))
df <- df0 %>% filter(catalog == "Kungliga")

@@

### Fig. 7: Publishing activity and riksdagar: Kungliga

```{r riksdar_kungliga, echo=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=5}

@@

@@

  riksplace <- c(rep("Unknown", 2), "Stockholm", "Stockholm", "Stockholm", "Norrköping", "Stockholm", rep("Unknown", 3), "Gävle", rep("Unknown", 7))
  #cols <- c("white", "Purple", "yellow", "orange", "darkgray")
  cols <- c("white", "gray", "darkgray")
  rectangles$place <- rectangles$shade
  rectangles$place[rectangles$place == "Highlight"] <- riksplace
  rectangles$place[rectangles$place == "Background"] <- "" 
  rectangles$shade <- factor(rectangles$shade)

  p <- p + geom_line(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + geom_point(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + scale_x_continuous(breaks=seq(min(dfm$date), max(dfm$date), 10))
  p <- p + ylab("Title count (n)")
  p <- p + xlab("Year")
  p <- p + ggtitle(paste(catal, " ", minyear, "-", maxyear, sep = ""))
  print(p)
```

@@

### Fig. 8: Publishing activity and riksdagar: Fennica

```{r riksdar_fennica, echo=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=5}
minyear <- 1760
maxyear <- 1860
catal <- "Fennica"

@@

### Fig 9A: Fennica: Publishers in Swedish University towns (and Finland)

```{r publishers2-finland, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- df0
catal <- "Fennica"
df <- df %>%
        filter(catalog == catal) %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

@@

theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack",
       		    aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +
       scale_fill_brewer(palette="Spectral") +       
       guides(fill = guide_legend(reverse = FALSE, title = "")) +
       ggtitle(catal)       
       #ggtitle(paste("Unique publishers in selected publication places (", catalogue, ")"))

print(p)

### Fig 9B: Kungliga: Publishers in Swedish University towns (and Finland)

```{r kungliga9b, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- df0
catal <- "Kungliga"
df <- df %>%
        filter(catalog == catal) %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

@@

theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack",
       		    aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +
       scale_fill_brewer(palette="Spectral") +       
       guides(fill = guide_legend(reverse = FALSE, title = "")) +
       ggtitle(catal)
       #ggtitle(paste("Unique publishers in selected publication places (", catalogue, ")"))

print(p)

@@

```{r publishers4-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=6, fig.width=12}
library(dplyr)
catalogue = "Fennica"
place <- "Vaasa"
minyear <- 1750
maxyear <- 1828

df <- df0 %>% filter(catalog == catalogue & publication_place == place) %>%	
        filter(publication_year >= minyear & publication_year <= maxyear)

p <- ggplot() + ggtitle("Fig 10: to be added")

# TITLE COUNT
#theme_set(theme_bw(20))
#p <- ggplot(npub, aes(x = publication_decade, y = publisher)) +
#      geom_point(aes(size = n)) + 
#      #geom_line(aes(color = publisher), size = 2) +
#      #geom_smooth(aes(fill = publisher)) +
#      # geom_bar(stat = "identity", position = "stack", aes(fill = publisher)) #+ 
#       xlab("Year") +
#       ylab("") +
#       guides(size = guide_legend(title = "Title count (n)"))
#       #ggtitle(paste("Title count per publisher (", place, "/", catalogue, ")"#, sep = ""))

@@

```{r publishers-fennica-11, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=6, fig.width=12}
library(dplyr)
place <- "Turku"
catalogue = "Fennica"
df = df0

@@

```{r percapita, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# Read historical population sizes
f <- system.file("extdata/population_sizes_in_cities.csv", package = "bibliographica")
pop <- read.csv(f)
pop <- melt(pop, "Year")
colnames(pop) <- c("publication_decade", "publication_place", "population")
pop <- as_data_frame(pop)
pop$publication_place <- as.character(pop$publication_place)

@@

### Fig. 13: Octavo paper consumption

@@

```{r octavo, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# Take cities from Kungliga except Turku
cat <- "Kungliga"
df1 <- df0 %>% filter(publication_place %in% setdiff(cities, "Turku")) %>%
      	       filter(catalog == cat & gatherings == "8vo") %>%
	       group_by(publication_decade, publication_place) %>%
	       summarise(n = n(), paper = sum(paper, na.rm = TRUE))

@@

### Fig. 14: Octavo title length (word count)

Average title length in words in Stockholm, Turku, Uppsala, Lund,
Göteborg and Linköping. According to Kungliga, except Turku from
Fennica. Only for octavo.

```{r octavo2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# Take cities from Kungliga except Turku
cat <- "Kungliga"
df1 <- df0 %>% filter(publication_place %in% setdiff(cities, "Turku")) %>%
      	       filter(catalog == cat & gatherings == "8vo") %>%
	       group_by(publication_decade, publication_place) %>%
	       summarise(n = n(), paper = sum(paper, na.rm = TRUE), 
	         mean_title_length = mean(title_length_wordcount, na.rm = TRUE))

# Take cities from Fennica for Turku
cat <- "Fennica"
df2 <- df0 %>% filter(publication_place %in% "Turku") %>%
      	       filter(catalog == cat & gatherings == "8vo") %>%
	       group_by(publication_decade, publication_place) %>%
	       summarise(n = n(), paper = sum(paper, na.rm = TRUE), 
                 mean_title_length = mean(title_length_wordcount, na.rm = TRUE))

@@

p <- ggplot(df, aes(x = publication_decade, y = mean_title_length)) +
       geom_line(aes(linetype = publication_place), size = 1) +
       geom_point(aes(shape = publication_place), size = 5) +
       xlab("Publication year") +
       ylab("Title length (words)") +
       ggtitle("Octavo: mean title length") +
       guides(shape = guide_legend(title = "Place")) +
       guides(linetype= guide_legend(title = "Place"))        
print(p)
```

@@

### Fig 15: Language use in Turku, Stockholm, Uppsala and Lund

```{r language-perc2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=12}
pics <- list()
theme_set(theme_bw(20))
myplaces <- c("Turku", "Greifswald", "Stockholm", "Uppsala", "Lund")

@@

### Fig. 16: Topic richness per language (Fennica)

```{r topics232, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
catalogue <- "Fennica"
df <- filter(df0, catalog == "Fennica") %>%
  filter(publication_year >= 1640 & publication_year <= 1828) %>%
  select(publication_year, language, subject_topic)
df <- unique(df)
df <- df %>% group_by(publication_year, language) %>%
             summarise(n = length(unique(unlist(strsplit(subject_topic, ";")))))

@@

### Fig. 17: Temporary sermons in Latin and Swedish

```{r topics882, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
  sel <- c("hautajaiset", "häät", "juhlamenot")
  catalogue <- "Fennica"
  df <- df0
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)

@@

### Fig. 18: Devotional literature (catechisms, hymns, prayers, etc.)

@@

```{r topics12122, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat")
catalogue <- "Fennica"

@@

### Fig. 19: Publisher per city over time

```{r publishersper, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
# Selected catalogue with selected decades
catalogue <- "Fennica"

df <- filter(df0, catalog == catalogue &
     		    publication_decade >= 1640 &
		    publication_decade <= 1828)

top.places <- names(top(df, "publication_place", 5))

df <- df %>%
      filter(publication_place %in% top.places) %>% 
      group_by(publication_decade, publication_place) %>%
      summarise(n = length(unique(na.omit(publisher))))

  theme_set(theme_bw(20))
  p <- ggplot(df, aes(x = publication_decade, y = n, shape = publication_place)) +
       geom_point(size = 5) +
       geom_line(size = 1, aes(linetype = publication_place)) +
       scale_color_brewer(palette="Greys") +
       #scale_fill_brewer(palette="Greys") +                     
       xlab("Publication year") +
       ylab("Publishers (n)") +
       guides(shape = guide_legend("Place")) +
       guides(linetype = guide_legend("Place")) +
       ggtitle(paste(catalogue))
  print(p)

>>>>

```{r 201606krakow-init2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Full combined catalogue (Fen + Kun) with marked duplicates
df.combined.preprocessed <- readRDS("df.combined.Rds")
# Data with duplicates removed
df.full <- df.combined.preprocessed %>% filter(!remove)
# Calculate title length (nchar and word count)
df.full$title_length_nchar <- nchar(df.full$title)
df.full$title_length_wordcount <- sapply(strsplit(df.full$title, " "), length)
df.full$title_length_wordcount[is.na(df.full$title)] <- NA

# Data with years limited as well
df0 <- df.full %>% filter(publication_year >= min.year & publication_year <= max.year)
```

@@

```{r Fig1, message=FALSE, warning=FALSE, fig.width=10, fig.height=4, echo=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw(20))
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>%
              group_by(catalog, publication_decade) %>%
	      summarise(n = length(unique(na.omit(publication_place))))
p <- ggplot(df, aes(x = publication_decade, y = n, shape = catalog, linetype = catalog)) + 
       geom_point(size = 4) + 
       geom_line(size = 1) + 
       ylab("Places (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen")) + 
       guides(shape = guide_legend(title = "Catalog", reverse = TRUE)) +
       guides(linetype = guide_legend(title = "Catalog", reverse = TRUE)) +       
       ggtitle("Unique publication places")
print(p)
```

@@

```{r Fig2, message=FALSE, warning=FALSE, fig.width=10, fig.height=4, echo=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw(20))
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>% group_by(catalog, publication_decade) %>% tally()
p <- ggplot(df, aes(x = publication_decade, y = n, shape = catalog, linetype = catalog)) + 
       geom_point(size = 4) + 
       geom_line(size = 1) + 
       ylab("Documents (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen")) +
       guides(shape = guide_legend(title = "Catalog", reverse = TRUE)) +
       guides(linetype = guide_legend(title = "Catalog", reverse = TRUE)) +  
       ggtitle("Title count")
print(p)
```

@@

```{r Fig3, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=9, fig.height=5, fig.show="hold", out.width="400px"}

for (catal in unique(df00$catalog)) {

  if (catal == "Kungliga") {
    df00 <- df.full %>% filter(publication_year >= min.year & publication_year <= max.year)
  } else if (catal == "Fennica") {
    df00 <- df.full %>% filter(publication_year >= min.year & publication_year <= 1900)
  }

  df2 <- subset(df00, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n())

  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 20)), "NA"))
  df2$highlight <- rep("Other", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "Octavo"
  df2$gatherings <- factor(df2$gatherings, levels = levels(df.full$gatherings))

@@

```{r Fig4, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
# Analysis time window
df <- df0 %>% filter(publication_year >= 1640 & publication_year <= 1911) %>%
              group_by(publication_decade) %>%
              summarize(mean_title_length = mean(title_length_wordcount),
	      		n = n())

for (catal in unique(df$catalog)) {
  dfs <- subset(df, catalog == catal)
  p <- ggplot(dfs, aes(x = publication_decade, y = mean_title_length)) +
       geom_point(aes(size = n), size = 5) +
       geom_line(size = 1) +
       ylab("Title length (word count)") +
       xlab("Publication decade") +
       ggtitle(paste("Title length (", catal, ")", sep = ""))

  print(p)
}
```

@@

```{r Fig5, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=18, fig.height=10}
pics <- list()
for (catalogue in na.omit(unique(df0$catalog))) {

@@

```{r Fig6, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=4}
library(magrittr)
library(ggplot2)
library(dplyr)
v <- seq(1500, 1800, 100)
theme_set(theme_bw(20))
df <- df0 %>% filter(catalog == "Kungliga")

@@

### Fig. 7A: Publishing activity and riksdagar: Kungliga

```{r Fig7A, echo=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=5}

@@

               1719-.5, 1719+.5, # Stockholm 20 januari 1719 1 juni 1719
               1720-.5, 1720+.5, # Stockholm 20 januari 1719 1 juni 1719
               1723-.5, 1723+.5, # Stockholm 20 januari 1719 1 juni 1719

               1726-.5, 1727+.5, # Stockholm 20 januari 1719 1 juni 1719

               1731-.5, 1731+.5, # Stockholm 20 januari 1719 1 juni 1719

               1734-.5, 1734+.5, # Stockholm 14 maj 1734 14 december 1734

               1738-.5, 1739+.5, #

               1740-.5, 1741+.5, #

               1742-.5, 1743+.5, #
               1746-.5, 1747+.5, #

               1751-.5, 1752+.5, #
               1755-.5, 1756+.5, #
               1760-.5, 1762+.5, #

               1765-.5, 1766+.5, # Stockholm 21 februari 1765 21 oktober 1766
               1769-.5, 1770+.5, # Norrköping & Stockholm 22 april 1769 5 februari 1770
               1771-.5, 1772+.5, # Stockholm 19 juni 1771 12 september 1772

@@

  riksplace <- c(rep("Unknown", 2), "Stockholm", "Stockholm", "Stockholm", "Norrköping", "Stockholm", rep("Unknown", 3), "Gävle", rep("Unknown", 7))
  #cols <- c("white", "Purple", "yellow", "orange", "darkgray")
  cols <- c("white", "lightgray", "gray", "darkgray")
  rectangles$shade[rectangles$shade == "Highlight"] <- rep(cols, length(which(rectangles$shade == "Highlight")))
  rectangles$shade <- factor(rectangles$shade)

  rectangles$place <- rectangles$shade
  rectangles$place[rectangles$place == "Highlight"] <- riksplace
  rectangles$place[rectangles$place == "Background"] <- ""

  # Draw Figure
  theme_set(theme_bw(20))
  p <- ggplot() +
         geom_rect(data = rectangles, 
           aes(xmin=xmin, xmax=xmax,
	       ymin=ymin, ymax=ymax,
	       fill=shade), alpha=0.8) +
         #scale_fill_manual(values = cols) +
         scale_fill_brewer(palette="Greys") +
	 guides(fill = "none")

  # p <- p + geom_text(data = rectangles, angle = 90, aes(x=xmin + .5, y=.9*ymax, label=place), size = 6)

  p <- p + geom_line(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + geom_point(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + scale_x_continuous(breaks=seq(min(dfm$date), max(dfm$date), 10))
  p <- p + ylab("Title count (n)")
  p <- p + xlab("Year")
  p <- p + ggtitle(paste(catal, " ", minyear, "-", maxyear, sep = ""))
  print(p)

@@

### Fig. 7B: Publishing activity and riksdagar: Fennica

```{r Fig7B, echo=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=5}
minyear <- 1760
maxyear <- 1860
catal <- "Fennica"

@@

### Fig 8: Publishers in Swedish University towns (Fennica)

Turku, Vaasa and Vyborg from Fennica; Stockholm, Lund, Uppsala, Greifswald, Tartu from Kungliga.

```{r Fig8, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- df0
catal <- "Fennica"
df <- df %>%
        filter(catalog == catal) %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

@@

theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack",
       		    aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +
       scale_fill_brewer(palette="Greys") +       
       guides(fill = guide_legend(reverse = FALSE, title = "")) +
       ggtitle(catal)       
       #ggtitle(paste("Unique publishers in selected publication places (", catalogue, ")"))

### Fig 9: Publishers in Swedish University towns (Kungliga)

All towns from Kungliga.

```{r Fig9, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- df0
catal <- "Kungliga"
df <- df %>%
        filter(catalog == catal) %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

@@

theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack",
       		    aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +
       scale_fill_brewer(palette="Greys") +       
       guides(fill = guide_legend(reverse = FALSE, title = "")) +
       ggtitle(catal)
       #ggtitle(paste("Unique publishers in selected publication places (", catalogue, ")"))

@@

```{r Fig10, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=6, fig.width=12}
library(dplyr)
catalogue <- "Fennica"
place <- "Vaasa"
minyear <- 1750
maxyear <- 1828

df <- df0 %>% filter(catalog == catalogue & publication_place == place) %>%	
        filter(publication_year >= minyear & publication_year <= maxyear) %>%
	group_by(publication_year) %>% 
	tally()

p <- ggplot(df, aes(x = publication_year, y = n)) +
     geom_bar(stat = "identity") +
     ggtitle("") +
     xlab("Publication year") +
     ylab("Title count")

@@

```{r Fig11, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=6, fig.width=12}
library(dplyr)
place <- "Turku"
catalogue = "Fennica"
df = df0

@@

```{r Fig12, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# Read historical population sizes
f <- system.file("extdata/population_sizes_in_cities.csv", package = "bibliographica")
pop <- read.csv(f)
pop <- melt(pop, "Year")
colnames(pop) <- c("publication_decade", "publication_place", "population")
pop <- as_data_frame(pop)
pop$publication_place <- as.character(pop$publication_place)

@@

### Fig. 13: Title count (absolute)

```{r Fig13, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
minyear <- 1640
maxyear <- 1828

p <- ggplot(df, aes(x = publication_decade, y = n)) +
       geom_line(aes(linetype = publication_place), size = 1) +
       geom_point(aes(shape = publication_place), size = 5) +
       xlab("Publication decade") +
       ylab("Title count per capita (n)") +
       ggtitle("Title count per capita") +
       guides(shape = guide_legend(title = "Place")) +
       guides(linetype = guide_legend(title = "Place")) 
print(p)
```

### Fig. 14: Octavo paper consumption

@@

```{r Fig14, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# Take cities from Kungliga except Turku
cat <- "Kungliga"
df1 <- df0 %>% filter(publication_place %in% setdiff(cities, "Turku")) %>%
      	       filter(catalog == cat & gatherings == "8vo") %>%
	       group_by(publication_decade, publication_place) %>%
	       summarise(n = n(), paper = sum(paper, na.rm = TRUE))

@@

### Fig. 15: Octavo title length (word count)

Average title length in words in Stockholm, Turku, Uppsala, Lund,
Göteborg and Linköping. According to Kungliga, except Turku from
Fennica.

```{r Fig15, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# Take cities from Kungliga except Turku
cat <- "Kungliga"

df00 <- df.full %>% filter(publication_year >= 1640 & publication_year <= 1850)

df1 <- df00 %>% filter(publication_place %in% setdiff(cities, "Turku")) %>%
      	       filter(catalog == cat) %>%
	       group_by(publication_decade, publication_place) %>%
	       summarise(n = n(), paper = sum(paper, na.rm = TRUE), 
	         mean_title_length = mean(title_length_wordcount, na.rm = TRUE))

# Take cities from Fennica for Turku
cat <- "Fennica"
df2 <- df00 %>% filter(publication_place %in% "Turku") %>%
      	       filter(catalog == cat) %>%
	       group_by(publication_decade, publication_place) %>%
	       summarise(n = n(), paper = sum(paper, na.rm = TRUE), 
                 mean_title_length = mean(title_length_wordcount, na.rm = TRUE))

@@

df$publication_place[!df$publication_place == "Stockholm"] <- "Other"

p <- ggplot(df, aes(x = publication_decade, y = mean_title_length)) +
       geom_line(aes(linetype = publication_place), size = 1) +
       geom_point(aes(shape = publication_place), size = 5) +
       xlab("Publication year") +
       ylab("Title length (words)") +
       ggtitle("Mean title length") +
       guides(shape = guide_legend(title = "Place")) +
       guides(linetype= guide_legend(title = "Place"))        
print(p)
```

@@

### Fig 16: Language use in Turku, Stockholm, Uppsala and Lund

```{r Fig16, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=12}
pics <- list()
theme_set(theme_bw(20))
myplaces <- c("Turku", "Greifswald", "Stockholm", "Uppsala", "Lund")

@@

### Fig. 17: Topic richness per language (Fennica)

```{r Fig17, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
catalogue <- "Fennica"
df <- filter(df0, catalog == "Fennica") %>%
  filter(publication_year >= 1640 & publication_year <= 1828) %>%
  select(publication_year, language, subject_topic)
df <- unique(df)
df <- df %>% group_by(publication_year, language) %>%
             summarise(n = length(unique(unlist(strsplit(subject_topic, ";")))))

@@

### Fig. 18: Temporary sermons in Latin and Swedish

```{r Fig18, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
  sel <- c("hautajaiset", "häät", "juhlamenot")
  catalogue <- "Fennica"
  df <- df0
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)

@@

### Fig. 19: Devotional literature (catechisms, hymns, prayers, etc.)

@@

```{r Fig19, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat")
catalogue <- "Fennica"

@@
